---
title: "Comment on Pruitt et al (2012) 'Behavioral Types of Predator and Prey Jointly Determine Prey Survival: Potential Implications for the Maintenance of Within-Species Behavioral Variation'"
author: "Sorbus torminalis"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_md: true
    toc: false

bibliography: bib/bib.bib
csl: bib/the-american-naturalist.csl   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, autodep = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(patchwork)
library(simba)  
source(here("pruitt", "sim.R"))# modified version of simba::sim to return both upper and lower triangle of distance matrix

set.seed(314)
theme_set(theme_bw())
```


```{r download}
#create data directory if needed
if(!fs::dir_exists(here("data"))){
  fs::dir_create(here("data"))
}

target_file <-  "Pruitt et al. Dryad File_AmNat2012.xls"
#download data if needed
if(!fs::file_exists(here("data", target_file))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.190pk253")
  fs::file_move(
    path = filelocation[[1]][1], 
    new_path = here("data", target_file)
  ) 
}
```


```{r import, message = FALSE}
#readxl::excel_sheets(here("data", target_file))

raw_data <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "raw data") %>% 
  clean_names() %>% 
  filter(!is.na(mesocosm)) %>% #remove mean & SD calculations
  rename(size = n_max)
  
repeatable <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "repeatable") %>% 
  select(1:4) %>% #junk in later columns
  clean_names()

shell_size_fear <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "shell size by fear") %>% 
  clean_names()

starfish <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "Starfish Attributes and Sel G", 
  skip = 1) %>% 
  clean_names()

#number of permutations for Ruggiero similarity null 
n_ruggiero_rep <- 1000
```

# Introduction

Several papers co-authored by Professor Pruitt have been shown to contain unexplained duplication in the raw data (e.g., @Laskowski2020-retract).
This comment reports an attempt to find duplications in @Pruitt2012.

```{r n_snails}
#range of number of snails per mesocosm
n_snails <- raw_data %>% 
  count(mesocosm) %>% 
  summarise(min = min(n), mx = max(n))

```


# Data description and methods

Partial raw data for @Pruitt2012 are archived as an Excel file [@Pruitt2011_data].
The sheet `raw data` includes data for activity, size (as `N Max`), anti-predator response, and survivorship for `r nrow(raw_data)` snails in `r max(raw_data$mesocosm)` mesocosms.
Size and anti-predator response are presented as _z_-scores.
Within each mesocosm, which have between `r n_snails$min` and `r n_snails$mx` snails, the data are sorted by survival and partially sorted by snail size (fig. \@ref(fig:raw-data-plot)).

```{r raw-data-plot, fig.cap = "Raw data from mesocosm 35. Blue and yellow points indicate snails that were eaten or survived respectively."}
#example of raw data
raw_data %>% 
  select(mesocosm, survival_1_0, activity_level, size, n_anti_pred_response) %>% 
  mutate(row = 1:n()) %>%
  filter(mesocosm == 35) %>% 
  pivot_longer(c(activity_level, size, n_anti_pred_response)) %>% 
  ggplot(aes(x = row, y = value, colour = factor(survival_1_0))) +
  geom_point(show.legend = FALSE) +
  scale_colour_viridis_d(option = "C", end = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  labs(x = "Row", y = "Value", colour = "Survival")
```

Because the data have been sorted, tests for duplicate sequences have a reduced utility because sequences may have been disrupted by sorting, or false positive duplicate sequences generated.

Instead, I test whether the overlap between measurements for the different mesocosms is higher than expected. 
With a slightly modified version of the `simba` package [@simba] in R [@R], I calculate the Ruggiero similarity [@Ruggiero1998] between mesocosms. 
The Ruggiero similarity ($a/(a + c)$) is one when all members of mesocosm 2 are in mesocosm 1. 
I make all duplicates values distinct, so if there are duplicate values in mesocosm 2, there needs to be at least as many duplicates of that value in mesocosm 1 to get a similarity of one. 

The expected distribution of Ruggiero similarities is estimated by permuting the data and recalculating the similarities `r n_ruggiero_rep` times.

Ruggiero similarities are less likely to be lower as diversity in the measurement increases. 
Table \@ref(tab:cardinality) shows that the anti-predator response has the lowest cardinality and diversity, whereas these are much higher for size, an important response variable. 

```{r cardinality}
#diversity of measurements
diversity <- raw_data %>% 
  select(size, n_anti_pred_response, activity_level) %>% 
  pivot_longer(everything(), names_to = "Variable") %>% 
  count(Variable, value) %>% 
  group_by(Variable) %>% 
  nest() %>% 
  mutate(
    `Shannon's H` = map_dbl(data, ~{select(.x, n) %>% vegan::diversity(index = "shannon", MARGIN = 2)}),
    N = map_int(data, nrow),
    `Hill's N1` = exp(`Shannon's H`)
  ) %>% 
  relocate(N, .after = Variable) %>% 
  select(-data)

diversity %>% 
  knitr::kable(digits = 2, 
               caption = "Cardinality (N), Shannon's index, and Hill's N1 diversity of the measurements for three snail response variables")
```



```{r ruggiero}
# similarity between size measurements in different mesocosms
ruggiero <- raw_data %>% 
  group_by(mesocosm, size) %>% 
  mutate(size2 = paste(size, 1:n(), sep = "_")) %>% 
  ungroup() %>% 
  count(mesocosm, size2) %>% 
  sim(method = "ruggiero", listout = FALSE, listin = TRUE) 

#convert distance matrix to long tibble
 ruggiero2 <- ruggiero %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "xid") %>% 
    pivot_longer(-xid, names_to = "yid", values_to = "rug") %>% 
    filter(xid != yid) %>%  # remove diagonal
    mutate(across(everything(), as.numeric))
 
#find similarity of mesocosms with permuted data 
random_ruggiero <- rerun(n_ruggiero_rep, {
  ruggiero <- raw_data %>% 
    mutate(size = sample(size)) %>% #randomise
    group_by(mesocosm, size) %>% 
    mutate(size2 = paste(size, 1:n(), sep = "_")) %>% 
    ungroup() %>% 
    count(mesocosm, size2) %>% 
    sim(method = "ruggiero", listout = FALSE, listin = TRUE) 

 ruggiero %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "xid") %>% 
    pivot_longer(-xid, names_to = "yid", values_to = "rug") %>% 
    filter(xid != yid) %>%  # remove diagonal
    mutate(across(everything(), as.numeric))
})

#find max of random trials
rugg_rand_max <- random_ruggiero %>% 
  set_names(1:length(.)) %>% 
  bind_rows(.id = "n") %>% 
  group_by(n) %>% 
  summarise(mx = max(rug), .groups = "drop") 

```

# Results


Figure \@ref(fig:ruggiero-plots) shows that the Ruggiero similarities between some pairs of mesocosms are much higher than expected from a random permutation of the data.
For example, `r sum(ruggiero2$rug == 1)` pairs of mesocosms have a Ruggiero similarity of 1 (fig. \@ref(fig:ruggiero-plots)A), that is the measurements in the smaller mesocosm are a subset of those in the larger mesocosm (fig. \@ref(fig:paired-plots)).
A further `r sum(ruggiero2$rug > 0.8 & ruggiero2$rug < 1)` pairs of mesocosms have a Ruggiero similarity above 0.8 (fig. \@ref(fig:ruggiero-plots)A).
In contrast, the 99^th^ percentile of the maximum similarity found in `r n_ruggiero_rep` permutations is `r quantile(rugg_rand_max$mx, prob = 0.99) %>% round(3)` and the highest maximum observed was `r max(rugg_rand_max$mx) %>% round(3)` (fig. \@ref(fig:ruggiero-plots)B).
Similarities above 0.8 are therefore expected to be very rare.

```{r ruggiero-plots, fig.cap = glue::glue("For the snail size data, A) maximum Ruggiero similiarity between mesocosms for each of {n_ruggiero_rep} permutations, B) distribution of Ruggiero similiarities between the mesocosms in the archived data, C) Ruggiero similiarities between pairs of mesocosms.")}

#raw data histogram
rugg_size_hist <- ruggiero2 %>% 
  ggplot(aes(x = rug)) + 
  geom_histogram(bins = 30) +
  labs(x = "Ruggiero similarity")
 
#random max similarity
rugg_rand_hist <- rugg_rand_max %>% 
  ggplot(aes(x = mx)) +
  geom_histogram(bins = 20) +
  labs(x = "Maximum Ruggiero similarity")

#raster
rugg_raster <- ruggiero2 %>% 
    ggplot(aes(x = xid, y = yid, fill = rug)) + 
    geom_raster() +
    scale_fill_viridis_c()  +
    scale_x_continuous(expand = c(0, 0), breaks = seq(2, 54, 2)) +
    scale_y_continuous(expand = c(0, 0), breaks = seq(2, 54, 2)) +
    labs(x = "Mesocosm 1", y = "Mesocosm 2", fill = "Ruggiero\nsimilarity") +
  theme(panel.grid = element_blank())

#compile plots with patchwork
(rugg_size_hist + rugg_rand_hist) / rugg_raster +
  patchwork::plot_annotation(tag_levels = "A") + plot_layout(heights = c(0.3, 0.7))
```

```{r paired-plots, fig.cap = "Size against row number for pairs of mesocosms which Ruggiero similarity of 1. Green points are in the paired mesocosm, blue points are not." }

# ruggiero2 %>% 
#   filter(rug  > 0.9) %>% 
#   rename(`Mesococsm 1` = xid, `Mesococsm 2` = yid, "Ruggiero similarity" = rug)

#look at pairs of mesocosms
#in a table
a <- raw_data %>% 
  filter(mesocosm %in% c(3, 4)) %>% 
  select(mesocosm, size) %>% 
  group_by(mesocosm) %>% 
  mutate(n = row_number(),
         size = sort(size)) %>% 
  pivot_wider(names_from = mesocosm, values_from = size)
#a
#graphically
plot_2_mecososms <- function(data = raw_data, m1, m2){
  mesos <- raw_data %>% 
    select(mesocosm, size) %>% 
    mutate(n = 1:nrow(.)) %>% 
    filter(mesocosm %in% c(m1, m2)) %>% 
    group_by(mesocosm, size) %>% 
    mutate(
      size2  = paste(size, 1:n(), sep = "_"), #make duplicates distinct
      alt = if_else(mesocosm == m1, true = m2, false = m1)
    )
  
  mesos %>% left_join(
    mesos %>% ungroup() %>% select(-mesocosm, -size), 
    by = c("mesocosm" = "alt", "size2" = "size2")) %>% 
    select(-alt) %>%
    mutate(mesocosm = factor(mesocosm, levels = c(m1, m2))) %>% 
    ggplot(aes(x = n.x, y = size, colour = !is.na(n.y))) + 
    geom_point(show.legend = FALSE) +
    scale_colour_viridis_d(end = 0.8) +
    facet_wrap(~mesocosm, scales = "free_x", ncol = 1)
}

meso_plots <- ruggiero2 %>% 
  filter(rug == 1) %>% 
  rowwise() %>% 
  mutate(plots = map2(.x = xid, .y = yid, ~plot_2_mecososms(m1 = .x, m2 = .y)))

wrap_plots(meso_plots$plots) + 
  plot_layout(guides = "collect")  +
  patchwork::plot_annotation(tag_levels = "A") & 
  theme(axis.title.x = element_blank())

# meso_plots <- ruggiero2 %>%
# filter(rug < 1 & rug > 0.8) %>%
# rowwise() %>%
# mutate(plots = map2(.x = xid, .y = yid, ~plot_2_mecososms(m1 = .x, m2 = .y)))
# wrap_plots(meso_plots$plots) +
# plot_layout(guides = "collect") &
# theme(axis.title.x = element_blank())

```

High similarities were not detected in either the activity or anti-predator response data (results not shown).


# Conclusions

The presence of unexpectedly high similarities between size measurements in the different mesocosms may be indicative of data duplication.
Because the data were sorted, the mesocosms are a convenient sized chunk of data to test for duplication, but if give reduced similarities if the duplication was done on a different scale.
Tests of duplication on the original unsorted data will be much more powerful.

Snail size is an important variable in @Pruitt2012.
If the size data cannot be relied upon, a large part of the results become questionable.

# Reproducibility

All analyses in this comment can be reproduced with code archived at https://github.com/Sorbus-torminalis/start/tree/main/pruitt


# References {-}
