---
title: "Comment on Pruitt et al (2012) 'Behavioral Types of Predator and Prey Jointly Determine Prey Survival: Potential Implications for the Maintenance of Within-Species Behavioral Variation'"
author: "Sorbus torminalis"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_md: true
    toc: false

bibliography: bib/bib.bib
csl: bib/the-american-naturalist.csl   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, autodep = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(patchwork)
library(simba)  
source(here("pruitt", "sim.R"))# modified version of simba::sim to return both upper and lower triangle of distance matrix

set.seed(314)
theme_set(theme_bw())
```


```{r download}
#create data directory if needed
if(!fs::dir_exists(here("data"))){
  fs::dir_create(here("data"))
}

target_file <-  "Pruitt et al. Dryad File_AmNat2012.xls"
#download data if needed
if(!fs::file_exists(here("data", target_file))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.190pk253")
  fs::file_move(
    path = filelocation[[1]][1], 
    new_path = here("data", target_file)
  ) 
}
```


```{r import, message = FALSE}
#readxl::excel_sheets(here("data", target_file))

raw_data <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "raw data") %>% 
  clean_names() %>% 
  filter(!is.na(mesocosm)) %>% #remove mean & SD calculations
  rename(size = n_max)
  
repeatable <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "repeatable") %>% 
  select(1:4) %>% #junk in later columns
  clean_names()

shell_size_fear <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "shell size by fear") %>% 
  clean_names()

starfish <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "Starfish Attributes and Sel G", 
  skip = 1) %>% 
  clean_names()

#number of permutations for Ruggiero similarity null 
n_ruggiero_rep <- 1000
```

# Introduction

Several papers co-authored by Professor Pruitt have been shown to contain unexplained duplication in the raw data (e.g., @Laskowski2020-retract).
This comment reports an attempt to find duplications in @Pruitt2012.

```{r n_snails}
#range of number of snails per mesocosm
n_snails <- raw_data %>% 
  count(mesocosm) %>% 
  summarise(min = min(n), mx = max(n), low  = sum(n < 37), n_meso = n())
```


# Data description and methods

Partial raw data for @Pruitt2012 are archived as an Excel file [@Pruitt2011_data].
The sheet `raw data` includes data for activity, size (as `N Max`), anti-predator response, and survivorship for `r nrow(raw_data)` snails in `r max(raw_data$mesocosm)` mesocosms.
Size and anti-predator response are presented as _z_-scores.
Mesocosms have between `r n_snails$min` and `r n_snails$mx` snails, whereas the paper reports that 37--46 snails are used in each mesocosm: `r n_snails$low` of the `r n_snails$n_meso` mesocosms have fewer snails than the reported minimum.
The number of snails per mesocosm is varied in replicates A and B, but consistent in replicate C (Fig. \@ref(fig:n-snail-plot)).

```{r n-snail-plot, fig.cap = "Number of snails per mesocosm.", fig.height = 4, fig.width = 6}
ggplot(raw_data, aes(x = factor(mesocosm))) + 
  geom_bar() + 
  facet_wrap(~replicate, scales = "free_x") +
  labs(x = "Mesocosm", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```


Within each mesocosm the data are sorted by survival and partially sorted by snail size (fig. \@ref(fig:raw-data-plot)).

```{r raw-data-plot, fig.cap = "Raw data from mesocosm 35. Blue and yellow points indicate snails that were eaten or survived respectively."}
#example of raw data
raw_data %>% 
  select(mesocosm, survival_1_0, activity_level, size, n_anti_pred_response) %>% 
  mutate(row = 1:n()) %>%
  filter(mesocosm == 35) %>% 
  pivot_longer(c(activity_level, size, n_anti_pred_response)) %>% 
  ggplot(aes(x = row, y = value, colour = factor(survival_1_0))) +
  geom_point(show.legend = FALSE) +
  scale_colour_viridis_d(option = "C", end = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  labs(x = "Row", y = "Value", colour = "Survival")
```

Because the data have been sorted, tests for duplicate sequences have a reduced utility because sequences may have been disrupted by sorting, or false positive duplicate sequences generated.

Instead, I test whether the overlap between measurements for the different mesocosms is higher than expected. 
With a slightly modified version of the `simba` package [@simba] in R [@R], I calculate the Ruggiero similarity [@Ruggiero1998] between mesocosms. 
The Ruggiero similarity ($a/(a + c)$) is one when all members of mesocosm 2 are in mesocosm 1. 
I make all duplicates values distinct, so if there are duplicate values in mesocosm 2, there needs to be at least as many duplicates of that value in mesocosm 1 to get a similarity of one. 

The expected distribution of Ruggiero similarities is estimated by permuting the snail sizes recalculating the similarities `r n_ruggiero_rep` times.
The paper reports that  "Randomly selected, individually marked snails ...  were  placed  within  each ... mesocosms".
Obviously the randomisation would be done within each replicate as they were gathered at different times (see appendix A), however, I found that permuting the measurements within each replicate gave slightly smaller maximum similarities than permuting across the entire dataset. 
I chose  to permute across the entire dataset as it is more favourable to the original authors.

Ruggiero similarities are less likely to be lower as diversity in the measurement increases. 
Table \@ref(tab:cardinality) shows that the anti-predator response has the lowest cardinality and diversity, whereas these are much higher for size, an important response variable. 

```{r cardinality}
#diversity of measurements
diversity <- raw_data %>% 
  select(size, n_anti_pred_response, activity_level) %>% 
  pivot_longer(everything(), names_to = "Variable") %>% 
  count(Variable, value) %>% 
  group_by(Variable) %>% 
  nest() %>% 
  mutate(
    `Shannon's H` = map_dbl(data, ~{select(.x, n) %>% vegan::diversity(index = "shannon", MARGIN = 2)}),
    N = map_int(data, nrow),
    `Hill's N1` = exp(`Shannon's H`)
  ) %>% 
  relocate(N, .after = Variable) %>% 
  select(-data)

diversity %>% 
  knitr::kable(digits = 2, 
               caption = "Cardinality (N), Shannon's index, and Hill's N1 diversity of the measurements for three snail response variables")
```



```{r ruggiero}
# similarity between size measurements in different mesocosms
ruggiero <- raw_data %>% 
  group_by(mesocosm, size) %>% 
  mutate(size2 = paste(size, 1:n(), sep = "_")) %>% 
  ungroup() %>% 
  count(mesocosm, size2) %>% 
  sim(method = "ruggiero", listout = FALSE, listin = TRUE) 

#convert distance matrix to long tibble
 ruggiero2 <- ruggiero %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "xid") %>% 
    pivot_longer(-xid, names_to = "yid", values_to = "rug") %>% 
    filter(xid != yid) %>%  # remove diagonal
    mutate(across(everything(), as.numeric))
 
#find similarity of mesocosms with permuted data 
random_ruggiero <- rerun(n_ruggiero_rep, {
  ruggiero <- raw_data %>%
    group_by(replicate) %>% 
    mutate(size = sample(size)) %>% #randomise
    group_by(mesocosm, size) %>% 
    mutate(size2 = paste(size, 1:n(), sep = "_")) %>% 
    ungroup() %>% 
    count(mesocosm, size2) %>% 
    sim(method = "ruggiero", listout = FALSE, listin = TRUE) 

 ruggiero %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "xid") %>% 
    pivot_longer(-xid, names_to = "yid", values_to = "rug") %>% 
    filter(xid != yid) %>%  # remove diagonal
    mutate(across(everything(), as.numeric))
})

#find max of random trials
rugg_rand_max <- random_ruggiero %>% 
  set_names(1:length(.)) %>% 
  bind_rows(.id = "n") %>% 
  group_by(n) %>% 
  summarise(mx = max(rug), .groups = "drop") 

#find similarity of mesocosms with permuted data per replicate
random_ruggiero_replicate <- rerun(n_ruggiero_rep, {
  ruggiero <- raw_data %>% 
    mutate(size = sample(size)) %>% #randomise
    group_by(mesocosm, size) %>% 
    mutate(size2 = paste(size, 1:n(), sep = "_")) %>% 
    ungroup() %>% 
    count(mesocosm, size2) %>% 
    sim(method = "ruggiero", listout = FALSE, listin = TRUE) 

 ruggiero %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "xid") %>% 
    pivot_longer(-xid, names_to = "yid", values_to = "rug") %>% 
    filter(xid != yid) %>%  # remove diagonal
    mutate(across(everything(), as.numeric))
})

#find max of random trials
rugg_rand_max_replicate <- random_ruggiero_replicate %>% 
  set_names(1:length(.)) %>% 
  bind_rows(.id = "n") %>% 
  group_by(n) %>% 
  summarise(mx = max(rug), .groups = "drop") 



```

```{r check-permutation-results}
stopifnot(quantile(rugg_rand_max$mx, prob = 0.95) > quantile(rugg_rand_max_replicate$mx, prob = 0.95))# stop if permute by replicate gives higher similarities

```


# Results


Figure \@ref(fig:ruggiero-plots) shows that the Ruggiero similarities between some pairs of mesocosms are much higher than expected from a random permutation of the data.
For example, `r sum(ruggiero2$rug == 1)` pairs of mesocosms have a Ruggiero similarity of 1 (fig. \@ref(fig:ruggiero-plots)A), that is the measurements in the smaller mesocosm are a subset of those in the larger mesocosm.
A further `r sum(ruggiero2$rug > 0.8 & ruggiero2$rug < 1)` pairs of mesocosms have a Ruggiero similarity above 0.8 (fig. \@ref(fig:ruggiero-plots)A).
In contrast, the 99^th^ percentile of the maximum similarity found in `r n_ruggiero_rep` permutations is `r quantile(rugg_rand_max$mx, prob = 0.99) %>% round(3)` and the highest maximum observed was `r max(rugg_rand_max$mx) %>% round(3)` (fig. \@ref(fig:ruggiero-plots)B).
Similarities between mesocosms above 0.8 are therefore expected to be very rare, and datasets with many such similarities should be exceedingly rare.

```{r replicate-C}
snail_meta <- raw_data %>% distinct(replicate, mesocosm)

ruggiero3 <- ruggiero2 %>%
  left_join(snail_meta, by = c(xid = "mesocosm")) %>% 
  mutate(xid = paste(replicate, formatC(xid, width = 2))) %>% 
  select(-replicate) %>% 
  left_join(snail_meta, by = c(yid = "mesocosm")) %>% 
  mutate(yid = paste(replicate, formatC(yid, width = 2))) %>% 
  select(-replicate) %>% 
  mutate(rug_class = case_when(rug  == 1 ~ "1",
                          rug > 0.9 ~ "> 0.9",
                          rug > 0.8 ~ "> 0.8",
                          rug > 0.7 ~ "> 0.7",
                          TRUE ~ "< 0.7")) 

C_max <- ruggiero3 %>% 
  group_by(xid) %>% filter(str_detect(xid, "C")) %>% 
  slice(which.max(rug)) %>% 
  ungroup() 
C_min <- C_max %>% 
  summarise(m = min(rug))
```


There are a few pairs of mesocosms with high similarities within replicates, but most of the pairs are when mesocsoms from replicates A and B are subsets of the larger replicate C mesocosms. 
All of the mesocosm in replicate C have a similarity of at least `r round(C_min$m, 3)` with mesocosms in replicates A or B.

High similarities were not detected in either the activity or anti-predator response data (results not shown).

```{r ruggiero-plots, fig.cap = glue::glue("For the snail size data, A) distribution of Ruggiero similiarities between the mesocosms in the archived data, B) maximum Ruggiero similiarity between mesocosms for each of {n_ruggiero_rep} permutations, C) Ruggiero similiarities between pairs of mesocosms (red lines separate replicates).")}

#raw data histogram
rugg_size_hist <- ruggiero2 %>% 
  ggplot(aes(x = rug)) + 
  geom_histogram(bins = 30) +
  labs(x = "Ruggiero similarity")
 
#random max similarity
rugg_rand_hist <- rugg_rand_max %>% 
  ggplot(aes(x = mx)) +
  geom_histogram(bins = 20) +
  labs(x = "Maximum Ruggiero similarity")

#raster

rugg_raster <- ruggiero3 %>% 
  ggplot(aes(x = xid, y = yid, fill = rug_class)) + 
    geom_raster() +
    scale_fill_viridis_d()  + 
    geom_vline(xintercept = c(18.5, 36.5), colour = "red") +
    geom_hline(yintercept = c(18.5, 36.5), colour = "red") +
    labs(x = "Mesocosm 1", y = "Mesocosm 2", fill = "Ruggiero\nsimilarity") +
    theme(panel.grid = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5))

#compile plots with patchwork
(rugg_size_hist + rugg_rand_hist) / rugg_raster +
  patchwork::plot_annotation(tag_levels = "A") + plot_layout(heights = c(0.3, 0.7))
```

## Overlap between replicate C and the other replicates

This section explores the high similarity between mesocosms in replicate C and the other two replicates in more detail.
Figure \@ref(fig:replicates-A-B-vs-C) shows for each which measurements in replicate C (one column per mesocosm) are also present in each mesocosm in replicates A and B (rows, arranged by mesocosm number).
Some measurements are expected to in common between the pairs of mesocosms, but long sequences are not.

Mesocosm 3 includes all measurements from mesocosm 4, measurements not in mesocosm 4 are in mesocosm 2. 
Almost all the measurement from mesocosm 6 overlap with measurements from mesocosm 5.
Further across, about half of mescosm 12 overlaps with mesocosm 8, and half with mesocosm 10, which in turn overlaps the first part of mesocosm 15.
Similar overlaps of whole or part mesocosms appear for most or all of replicate C.
This figure is consistent with the size measurements for replicate C being a duplicate of measurements in the other replicates.

Equivalent plots for within replicates A and B and between replicates A and B are in appendix B where there are fewer overlaps, but still several pairs of mesocosms with higher similarity than expected.



```{r replicates-A-B-vs-C, fig.height = 9, fig.width = 9.5, fig.cap = "Measurements in common (purple) between mesocosms in replicate C (columns) and mesocosms in replicates A and B (rows). Axes are size against snail row number within replicate C."}
#extract replicate C mesocosms
replicate_C <- raw_data %>% 
  filter(replicate == "C") %>% 
  mutate(n = 1:n()) %>% #sequential numbers
  select(-starfish_id, -replicate, mesocosmC = mesocosm, -activity_level, -n_anti_pred_response)

raw_data %>% 
  filter(replicate != "C") %>% #get other replicates
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, # for each mesocosm in A/B find matching measurements in replicate C
                  ~left_join(
                    replicate_C, 
                    .x %>% mutate(one = 1), #dummy column to show match made
                    by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
  scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmC, scales = "free_x") +
  theme(panel.spacing = unit(0.4, "mm"), strip.text.y = element_text(angle = 0))

```






# Conclusions

The presence of unexpectedly high similarities between size measurements in the different mesocosms may be indicative of data duplication.
Tests of duplication on the original unsorted data will be much more powerful.

Snail size is an important variable in @Pruitt2012.
If it cannot be relied upon, a large part of the results become questionable.

# Reproducibility

All analyses in this comment can be reproduced with code archived at https://github.com/Sorbus-torminalis/start/tree/main/pruitt


# References {-}


<div id="refs"></div>

# Appendix A {-}

Anova results of size against mesocosm for each replicate analysed separately. 
All three anova are highly significant, suggesting that the randomisation was not very effective. 
Figure \@ref(fig:mesocosm-size-plots) shows that some mesocosms in each replicate have smaller sized snails that expected. 
Given the immense challenge of randomising nearly a thousand snails identified by paint spots, I suspect that the snails were haphazardly assigned to mesocosms and that small snails tended to be picked last.
I don't know what, if any, effect this will have on the original result.
It will make the null model used here give lower maximum similarities than one that included the incomplete randomisation.
I did not attempt to quantify this effect as I expect it to be small.

## Replicate A {-}

```{r mesocosm-size-models-A}
mod_A <- raw_data %>% 
  filter(replicate == "A") %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
  lm(size ~ mesocosm, data = .)

anova(mod_A)
```

## Replicate B {-}

```{r mesocosm-size-models-B}
mod_B <- raw_data %>% 
  filter(replicate == "B") %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
  lm(size ~ mesocosm, data = .)

anova(mod_B)
```

## Replicate C {-}

```{r mesocosm-size-models-C}
mod_C <- raw_data %>% 
  filter(replicate == "C") %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
  lm(size ~ mesocosm, data = .)

anova(mod_C)
```


```{r mesocosm-size-plots, fig.cap = "Snail size for each mesocosm in each replicate."}
raw_data %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
ggplot(aes(x = mesocosm, y = size, fill = replicate)) +
  geom_boxplot(show.legend = FALSE) +
  geom_hline(aes(yintercept = y), raw_data %>% 
               group_by(replicate) %>% 
               summarise(y = mean(size), .groups = "drop")) +
  facet_wrap(~ replicate, scales = "free_x")
```

# Appendix B {-}

## Within replicate A

```{r replicateA}
## within A
raw_data %>% 
  filter(replicate == "A") %>% 
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, ~left_join(raw_data %>% mutate(n = 1:nrow(.)) %>% 
  filter(replicate == "A") %>% 
  select(mesocosmA = mesocosm, size, n), .x %>% mutate(one = 1), by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  filter(mesocosm != mesocosmA) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
   scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmA, scales = "free_x") +
  theme(panel.spacing = unit(0.5, "mm"), strip.text.y = element_text(angle = 0))
```

## Within replicate B {-}

```{r replicateB}
## within B
raw_data %>% 
  filter(replicate == "B") %>% 
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, ~left_join(raw_data %>% mutate(n = 1:nrow(.)) %>% 
  filter(replicate == "B") %>% 
  select(mesocosmB = mesocosm, size, n), .x %>% mutate(one = 1), by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  filter(mesocosm != mesocosmB) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
   scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmB, scales = "free_x") +
  theme(panel.spacing = unit(0.5, "mm"), strip.text.y = element_text(angle = 0))
```


## Replicate A vs replicate B {-}

```{r replicates-A-vs-B}
## A vs B
raw_data %>% 
  filter(replicate == "A") %>% 
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, ~left_join(raw_data %>% mutate(n = 1:nrow(.)) %>% 
  filter(replicate == "B") %>% 
  select(mesocosmB = mesocosm, size, n), .x %>% mutate(one = 1), by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  filter(mesocosm != mesocosmB) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
   scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmB, scales = "free_x") +
  theme(panel.spacing = unit(0.5, "mm"), strip.text.y = element_text(angle = 0))
```
