---
title: "Post-publication review of 'Smaller and bolder prey snails have higher survival in staged encounters with the sea star _Pisaster giganteus_'"
author: "Sorbus torminalis"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(readxl)
library(tidyverse)
library(janitor)
library(lme4)
library(car)
library(assertr)
theme_set(theme_bw())
```

```{r load-data, include = FALSE, echo = FALSE}
f <- here("data", "Foster.etal.CurrentZoo2017_Original.xlsx")


single <- read_excel(f, sheet = "Single Density Survival Data") %>% 
  slice(-1) %>% 
  mutate(`Survived (1/0)...13` = as.numeric(`Survived (1/0)...13`),
         `Survived (1/0)...25` = as.numeric(`Survived (1/0)...25`)
         )

double <- read_excel(f, sheet = "Double Density Survival Data") %>% 
  slice(-1) 
         
         
snail_original <- bind_rows(
  single = single, 
  double = double, 
  .id = "Density"
) %>% 
  rename(trial_ID = `Trial ID...1`,         
         snail_ID = `Snail ID...2`,
         seastar_ID = `Sea Star ID...3`, 
         snail_width_mm = `Snail Width (mm)...4`, 
         survived_01 = `Survived (1/0)...13`) %>% 
  filter(!is.na(snail_ID)) %>% 
  verify(survived_01 == `Survived (1/0)...25`) %>% #check duplicate columns identical
  select(-matches("\\.{3}")) %>% 
  clean_names() %>% 
  mutate(survived = if_else(survived_01 == 1, "Alive", "Dead"), 
         survived = factor(survived)) 

# snails <- read_excel(here("data", "Foster.etal.CurrentZoo2017.xlsx"), sheet = "Survival Data") %>% 
#   mutate( Survived = if_else(Survived_1.0 == 1, "Alive", "Dead"))
```

```{r}
ggplot(snail_original, aes(x = trial_id, y = snail_width_mm, fill = trial_id) ) + 
  geom_violin(show.legend = FALSE) 

lm(snail_width_mm ~ trial_id, data = snail_original) %>% 
  anova() # highly significant

leveneTest(snail_width_mm ~ as.factor(trial_id), data = snail_original) # not significant
```

## Figure 1

Figure 1 cannot be replicated - outliers have been removed. 
Plotting without jitter misrepresents how many points are in the tails of the distribution.
(Not sure boxplots will work well on integer data with low variance)

```{r figure_1}
ggplot(snail_original, aes(x = survived, y = snail_width_mm)) +
  geom_boxplot(outlier.shape = NA, fill = "grey60") +
  geom_jitter(height = 0.1, aes(colour = density)) +
  coord_flip()

ggplot(snail_original, aes(x = snail_width_mm, y = survived_01, colour = seastar_id)) +
  geom_jitter(width = 0, height = 0.2) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 5), method.args = list(family = binomial)) +
  facet_wrap(~ trial_id)

```

## Figure 2 

Figure 2 cannot be replicated - outliers have been removed. 


```{r average-height-out-water}
ggplot(snail_original, aes(x = survived, y = average_height_cm)) +
  geom_boxplot(outlier.shape = NA, fill = "grey60") +
  geom_jitter(height = 0, aes(colour = density)) +
  coord_flip() +
  scale_x_discrete(limits = c("Dead", "Alive")) #to match inconsistency in published

```

```{r heights}
# mean height - no formula in excel sheet
snail_with_height <- snail_original %>% 
    mutate(mean_max_height = rowMeans(select(., matches("trial_\\d_max_height")))) 

snail_with_height %>% 
  ggplot(aes(x = average_height_cm, mean_max_height)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, colour = "red")

ggplot(snail_with_height, aes(x = survived, y = mean_max_height)) +
  geom_boxplot(outlier.shape = NA, fill = "grey60") +
  geom_jitter(height = 0, aes(colour = density)) +
  coord_flip() +
  scale_x_discrete(limits = c("Dead", "Alive")) #to match inconsistency in published

```



## Correlations
 
Correlations can be reproduced.

```{r correlations}
#reproduces
snail_original %>% 
  select(snail_width_mm, peak_height_cm, average_height_cm, average_time_out) %>% 
  cor() %>% 
  round(4)


```


## GLMM model


> We used a GLMM with a binomial error distribution and log–link function to evaluate selection on prey traits. 

Binomial with log-link is possible but unusual and often fails to converge. 
This model fails to converge for me with a log-link.
This may simply be a typo.

> We included the individuals’ average maximum height out of the water obtained across their 5 anti-predator assays (height above water), max shell diameter (shell diameter), and prey density treatment (density treatment: 25 or 50 snails) as predictor variables in our model. 

Seems to use average height (mean of two minute measurements over five trials) rather than average maximum height. 

> We used snail survival as our binary response variable and individual P. giganteus identity and replicate ID as random effects in our model. 

Reasonable.

```{r}
model <- glmer(
  survived ~ density + snail_width_mm + average_height_cm + (1|trial_id) + (1|seastar_id), 
  family = binomial(link = "logit"), # log link fails 
  data = snail_with_height)
summary(model)

model2 <- glmer(# with mean max height
  survived ~ density + snail_width_mm + mean_max_height + (1|trial_id) + (1|seastar_id), 
  family = binomial(link = "logit"),
  data = snail_with_height)
summary(model2)
```

```{r}
drop1(model, test = "Chisq")



```


## Repeat measurements

Repeat behaviour measures are interesting - correlation decays over time. 


```{r}
pca_height <- snail_original %>% select(matches("trial_\\d_max_height")) %>% 
  prcomp(scale. = FALSE) 

biplot(pca_height)
screeplot(pca_height, bstick = TRUE)

snail_original %>% select(matches("trial_\\d_max_height")) %>% cor()


pca_time <- snail_original %>% select(matches("trial_\\d_time_out")) %>% 
  prcomp(scale. = TRUE) 
biplot(pca_time)
screeplot(pca_time, bstick = TRUE)

snail_original %>% select(matches("trial_\\d_time_out")) %>% cor()

```

## Behaviour metrics

Good correlation between max height and time out of water.

```{r}
cor(snail_original$trial_1_max_height, snail_original$trial_1_time_out)
cor(snail_original$trial_2_max_height, snail_original$trial_2_time_out)
cor(snail_original$trial_3_max_height, snail_original$trial_3_time_out)
cor(snail_original$trial_4_max_height, snail_original$trial_4_time_out)

```

