---
title: "Post-publication review of 'Cannibalism by damselflies increases with rising temperature'"
author: "Sorbus torminalis"
date: "9/5/2020"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(broom)
library(patchwork)
theme_set(theme_bw())
```

```{r download}
if(!fs::file_exists(here("data", "cannibalism_damsel_warming_april10_2017.xlsx"))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.7f0c4")
  fs::file_move(
    path = filelocation[[1]], 
    new_path = here("data", "cannibalism_damsel_warming_april10_2017.xlsx")
  ) 
}
```

```{r import}
cannibals <- readxl::read_xlsx(
  path = here("data", "cannibalism_damsel_warming_april10_2017.xlsx"), 
  sheet = "Experiment1-matrix", na = "NA")
```


## Head width measurements

There are two sets of head width measurements on the pairs of damselfly larvae: the first taken at day 0, the second at day 44. 
ImageJ is used to measure the widths.
_Bombus Apollineus_ noted in #1 that only for the second pair of measurement can the size difference between the two larvae be readily recalculated from the archived data.
That is not the only difference between the two sets of measurements.

The first set of measurements is reported as integer values whereas the second is reported to two decimal places. 
```{r}
widths <- cannibals %>% 
  select(matches("^T\\d_Hw_\\d")) %>% 
  pivot_longer(everything(), names_pattern = "^(T\\d)_Hw_(\\d)$", names_to = c("Time", "Individual")) %>% 
  filter(!is.na(value))

last_digit <- widths %>% 
  mutate(
    value = round(value, 4), #drop any excel weirdness
    value = floor(value), # drop decimal places from T2
    last10 = value %% 10,
    last100 = value %% 100) %>%
  pivot_longer(starts_with("last"), names_to = "ten_hundred", values_to = "last_digits") %>% 
  group_by(Time, Individual, ten_hundred) %>%  
  count(last_digits)

```

A more curious difference is in the frequency of the last one and two digits in the integer part of the data. 
The second measurements (T2; n = `r sum(widths$Time == "T2")`) have an approximately uniform distribution of values, as expected from Benford's law on last digits.
The first measurements (T1; n = `r sum(widths$Time == "T1")`; n is larger as there was no mortality at this stage) have an uneven distribution.
While many of the values end in two, none end in three, and few end in six or seven.
Several of the most prominent spikes in the distribution of the last two digits are where both digits are the same. For example, 22 occurs `r sum(widths$Time == "T1" & widths$value %% 100 == 22)` times, and 99 occurs `r sum(widths$Time == "T1" & widths$value %% 100 == 99)` times.

```{r}
last_digit_plot <- last_digit %>%
  filter(ten_hundred == "last10") %>% 
  ggplot(aes(x = last_digits, y = n, fill = Individual)) +
  geom_col() +
  scale_x_continuous(breaks = 0:9) +
  labs(x = "Last digit") +
  facet_wrap(~ Time, scales = "free")

last2_digit_plot <- last_digit %>%
  filter(ten_hundred == "last100") %>% 
  ggplot(aes(x = last_digits, y = n, fill = Individual)) +
  geom_col() +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Last two digits") +
  facet_wrap(~ Time, scales = "free")


last_digit_plot / last2_digit_plot + plot_layout(guides = "collect", )
```

It is not clear how this pattern in the first measurements could be generated as insufficient details are given in the methods (camera, resolution, etc). 

If the image resolution is low, there might be some issues with aliasing. 
For example, if the resolution is 2 $\mu m$ per pixel, then only widths that are multiples of 2 $\mu m$ are possible. 
However, this only applies if the larvae are perfectly orientated as measurements made diagonally could still give odd values.
Moreover, it is difficult to see how this could affect only one of the sets of measurements that were presumably collected with similar methods.

An independent dataset of damselfly head width measurements (https://doi.org/10.5061/dryad.7d85g7b), also measured with imageJ, showed a near uniform (after converting to $\mu m$) distribution of last digits.

Do the authors have any explanation for this pattern in the first set of width measurements?

Code to reproduce these analyses can be found at [https://github.com/Sorbus-torminalis/start/blob/main/R/cannibalism.Rmd](https://github.com/Sorbus-torminalis/start/blob/main/R/cannibalism.Rmd)