---
title: "Post-publication review of 'Cannibalism by damselflies increases with rising temperature'"
author: "Sorbus torminalis"
date: "9/5/2020"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(broom)
library(patchwork)
theme_set(theme_bw())
```

```{r download}
if(!fs::file_exists(here("data", "cannibalism_damsel_warming_april10_2017.xlsx"))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.7f0c4")
  fs::file_move(
    path = filelocation[[1]], 
    new_path = here("data", "cannibalism_damsel_warming_april10_2017.xlsx")
  ) 
}
```

```{r import}
cannibals <- readxl::read_xlsx(
  path = here("data", "cannibalism_damsel_warming_april10_2017.xlsx"), 
  sheet = "Experiment1-matrix", na = "NA") %>% 
  filter(!is.na(label))
```


## Head width measurements

There are two sets of head width measurements on the pairs of damselfly larvae: the first taken at day 0, the second at day 44. 
ImageJ is used to measure the widths.
_Bombus Apollineus_ noted in #1 that only for the second pair of measurement can the size difference between the two larvae be readily recalculated from the archived data.
That is not the only difference between the two sets of measurements.

The first set of measurements is reported as integer values whereas the second is reported to two decimal places. 
```{r}
widths <- cannibals %>% 
  select(matches("^T\\d_Hw_\\d")) %>% 
  pivot_longer(everything(), names_pattern = "^(T\\d)_Hw_(\\d)$", names_to = c("Time", "Individual")) %>% 
  filter(!is.na(value))

last_digit <- widths %>% 
  mutate(
    value = round(value, 4), #drop any excel weirdness
    value = floor(value), # drop decimal places from T2
    last10 = value %% 10,
    last100 = value %% 100) %>%
  pivot_longer(starts_with("last"), names_to = "ten_hundred", values_to = "last_digits") %>% 
  group_by(Time, Individual, ten_hundred) %>%  
  count(last_digits)

```

A more curious difference is in the frequency of the last one and two digits in the integer part of the data. 
The second measurements (T2; n = `r sum(widths$Time == "T2")`) have an approximately uniform distribution of values, as expected from Benford's law on last digits.
The first measurements (T1; n = `r sum(widths$Time == "T1")`; n is larger as there was no mortality at this stage) have an uneven distribution.
While many of the values end in two, none end in three, and few end in six or seven.
Several of the most prominent spikes in the distribution of the last two digits are where both digits are the same. For example, 22 occurs `r sum(widths$Time == "T1" & widths$value %% 100 == 22)` times, and 99 occurs `r sum(widths$Time == "T1" & widths$value %% 100 == 99)` times.

```{r}
last_digit_plot <- last_digit %>%
  filter(ten_hundred == "last10") %>% 
  ggplot(aes(x = last_digits, y = n, fill = Individual)) +
  geom_col() +
  scale_x_continuous(breaks = 0:9) +
  labs(x = "Last digit") +
  facet_wrap(~ Time, scales = "free")

last2_digit_plot <- last_digit %>%
  filter(ten_hundred == "last100") %>% 
  ggplot(aes(x = last_digits, y = n, fill = Individual)) +
  geom_col() +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Last two digits") +
  facet_wrap(~ Time, scales = "free")


last_digit_plot / last2_digit_plot + plot_layout(guides = "collect", )
```

It is not clear how this pattern in the first measurements could be generated as insufficient details are given in the methods (camera, resolution, etc). 

If the image resolution is low, there might be some issues with aliasing. 
For example, if the resolution is 2 $\mu m$ per pixel, then only widths that are multiples of 2 $\mu m$ are possible. 
However, this only applies if the larvae are perfectly orientated as measurements made diagonally could still give odd values.
Moreover, it is difficult to see how this could affect only one of the sets of measurements that were presumably collected with similar methods.

An independent dataset of damselfly head width measurements (https://doi.org/10.5061/dryad.7d85g7b), also measured with imageJ, showed a near uniform (after converting to $\mu m$) distribution of last digits.

Do the authors have any explanation for this pattern in the first set of width measurements?

Code to reproduce these analyses can be found at [https://github.com/Sorbus-torminalis/start/blob/main/R/cannibalism.Rmd](https://github.com/Sorbus-torminalis/start/blob/main/R/cannibalism.Rmd)



## EXTRA - extended analysis of last digits

```{r, last-extended}
library(twopointzerothree)

## HW1 - last pair of digits
last2 <- cannibals %>% 
  select(matches("^T1_Hw_\\d")) %>% 
  pivot_longer(everything()) %>% 
  filter(!is.na(value)) %>% 
  pull(value) %>%
  last2_diff_test(10000)
last2$p

autoplot(last2)
```

## Round 2

```{r}
cannibals %>% ggplot(aes(x = Timing, y = T1_Hw_1 - T1_Hw_2, colour = factor(Warming))) + 
  #geom_point(position = position_dodge(width = 1)) + 
  stat_summary(fun.data ="mean_sdl", fun.args = list(mult = 1), position = position_dodge(width = 1))
cannibals %>% ggplot(aes(x = Timing, y = T1_Hw_diff, colour = factor(Warming))) + 
#  geom_point(position = position_dodge(width = 1)) + 
  stat_summary(fun.data ="mean_sdl", fun.args = list(mult = 1), position = position_dodge(width = 1))

cannibals %>% group_by(Warming, Timing) %>% 
  summarise(pre_calc_mean = mean(T1_Hw_diff, na.rm = TRUE), pre_calc_sd = sd(T1_Hw_diff, na.rm = TRUE),
            mean = mean(T1_Hw_1 - T1_Hw_2, na.rm = TRUE), sd = sd(T1_Hw_1 - T1_Hw_2, na.rm = TRUE))


```


There is another unexpected pattern in the head width data for T1.

For Timing = 0, the mean difference between the head widths is zero for both levels of warming (this is the difference calculate from the two head widths, not the pre-calculated difference).

If one takes the data for Timing = 0 and, within each warming treatment, sorts the head width data for each individual, the sorted values are identical.

The head width measurements for the second individual are identical to the permuted head width measurements of the first. 

Given that there are 15 replicates of each treatment, and values span a range of at least 60 units, it seems unlikely that this pattern would happen by chance.

Please can the authors explain what happened.

```{r echo = TRUE}
sorted_HW <- cannibals %>% 
  select(Timing, Warming, matches("T1_Hw_\\d")) %>%
  filter(Timing == 0) %>%
  group_by(Warming) %>% 
  transmute(T1_Hw_1_sort = sort(T1_Hw_1), T1_Hw_2_sort = sort(T1_Hw_2)) 

identical(sorted_HW$T1_Hw_1_sort, sorted_HW$T1_Hw_2_sort)

sorted_HW %>% pander::pander(style = "rmarkdown")
```

## Round 3

Figure 2b in this paper shows activity rate for the two temperature treatments.

The error bars are asymmetric, which is unexpected as the caption reports that "All error bars represent 1 s.d."


The sheet 4 of the archived excel file for this study includes a plot the same data, also with asymmetric error bars. 
The data elements of this plot appear to be identical to the published plot.

The excel plot is based on a small table, apparently copied from a pivot table, which has columns for the mean activity and the standard deviation, and columns labelled `1.5.st.dev.activity`	and `0.5st.dev` which are used for the upper and lower error bar, respectively.

Formula in the cells show that these are actually 0.8 * SD and 0.2 * SD respectively.

Thus, the plot in excel, which appears identical to the published plot, does not have error bars of +/- SD but of +0.8/-0.2 SD. 

I am not aware of any theoretical justification for this choice of error bar, and none is given in the paper.

Please can the authors explain why they calculate error bars in this way and state whether asymmetric error bars in other papers (including https://pubpeer.com/publications/6B420F47E530C253F39D53E7576490#6) are calculated in the same way.



