---
title: "Post-publication review of 'Sexual dimorphism in a top predator (Notophthalmus viridescens) drives aquatic prey community assembly'"
author: "Sorbus torminalis"
date: "9/12/2020"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(broom)
library(patchwork)
theme_set(theme_bw())
```

This paper reports the effect of the sex ratio of newts (either 9M : 16F or vice versa) on the composition of aquatic prey communities.

I have no comments to make about the data, but I do have concerns about some of the statistical analyses. 

# Over dispersion

The paper fits GLMs to the abundance of each taxon with a Poisson error distribution. 
The results of this analysis, shown in table S1, can be replicated almost perfectly.

```{r download}
if(!fs::file_exists(here("data", "rspb20181717supp2.xlsx"))){
  print("Downloading file")
download.file(
  url = "https://royalsocietypublishing.org/action/downloadSupplement?doi=10.1098%2Frspb.2018.1717&file=rspb20181717supp2.xlsx", 
  destfile = here("data", "rspb20181717supp2.xlsx")
  ) 
}
```

```{r import}
dimorph <- readxl::read_xlsx(
  path = here("data", "rspb20181717supp2.xlsx"), 
  sheet = "Sheet1", na = "NA") %>% 
  mutate(treatment = factor(treatment, levels = c("MB", "FB")))
```

```{r reproduce-tableS1, results = "asis"}
dimorph_long <- dimorph %>% 
  select(
    tank, treatment, chironomid, notonectidae, coleoptera_adult, coleoptera_larvae,   
    odonata, ephemeroptera, megaloptera, cyclopoid_copepod, calanoid_copepod,
    harpacticoid_copepod, ostracod, other_cladocera, chdorinid_cladocera, 
    rotifer, corixid) %>% 
  pivot_longer(-c(tank, treatment), names_to = "taxa", values_to = "count") %>% 
  # #merge cladocera
  # mutate(taxa = recode(taxa, "chdorinid_cladocera" = "other_cladocera")) %>%  
  # group_by(tank, treatment, taxa) %>% 
  # summarise(count = sum(count)) %>% 
  mutate(habitat = case_when(
    taxa %in% c("chironomid", "coleoptera_larvae", "odonata", "ephemeroptera", "megaloptera", "harpacticoid_copepod", "ostracod") ~ "Benthic",
    taxa %in% c("notonectidae", "cyclopoid_copepod", "calanoid_copepod", "rotifer",  "other_cladocera") ~ "Pelagic", 
    taxa %in% c("coleoptera_adult", "corixid") ~ "Both", 
    TRUE ~ "not in table")) %>% #chdorinid_cladocera
  group_by(habitat, taxa) %>% 
  nest() %>% 
  arrange(habitat, taxa)

tableS1 <- dimorph_long %>% 
  mutate(
    mod = map(data, ~glm(count ~ treatment, data = .x, family = poisson)),
    coef = map(mod, tidy),
    mod_q = map(data, ~glm(count ~ treatment, data = .x, family = quasipoisson)),
    coef_q = map(mod_q, tidy),
    pval_q = map_dbl(coef_q, ~pull(filter(.x, term == "treatmentFB"), p.value)),
    dispersion = map_dbl(mod_q, ~summary(.x)$dispersion)
    ) %>% 
  unnest(coef) %>% 
  filter(term == "treatmentFB") %>% 
  select(-mod, -mod_q, -coef_q, -term, -data, -std.error, -statistic) %>% 
  mutate(across(c(estimate, dispersion), round, digits = 2), 
         across(c(p.value, pval_q), round, digits = 3)) %>% 
  rename(p_poisson = p.value, p_quasi = pval_q)

tableS1 %>% 
  pander::pander(style = "rmarkdown", split.tables = Inf)

```

The coefficient for other_cladocera differs slightly (merging chdorinid_cladocera with other_cladocera does not help), and the signs of the coefficients for Corixid, calanoid_copepod and rotifers are positive in table S1 but negative in the reproduction.
Otherwise the results are mostly the similar.

The problem with this analysis is that the counts are over-dispersed, as ecological data tend to be, as indicated by the dispersion column.
One way to account for over-dispersion in count data is to use the quasipoisson error distribution instead of the Poisson.
If this is done, the evidence for an effect of newt sex ratio on the abundance of taxa is reduced. 
Four of the seven taxa that apparently had statistically significant relationships at the p = 0.05 level are not statistically significant with the quasipoisson is used.

```{r multiple-testing, results = "hide"}
tableS1 %>% ungroup() %>% summarise(n = sum(p_poisson < 0.05), nq = sum(p_quasi < 0.05))
tableS1$p_quasi %>% p.adjust(method = "fdr") %>% min()
```


A second issue is one of multiple testing. 
If the p-values were adjusted to correct for multiple testing, the apparent strength of the evidence would weaken further.

## Ordinations

Although I have tried several different combinations of species in the NMDS I have been unable to get a configuration close to that published. 
I can get a configuration where the male and female dominate tanks separate, but this configuration is not robust to the inclusion of motile taxa. 
If non-default options have been used in the NMDS, that could well explain the  differences between the published figure and what I can reproduce.

Please can the authors archive their code for the ordinations 

```{r, eval = FALSE}
library(vegan)
library(ggvegan)
spp <- dimorph %>% 
  select(-tank, -treatment, -total_B, -total_P, -prop_B, -unknown) 

spp <- spp %>% 
  select(-c(corixid, notonectidae, coleoptera_adult, odonata, ephemeroptera, megaloptera)) %>% #motile
  select(  -harpacticoid_copepod, -`other_benthic_(e.g.,_newt_larvae)`) %>% #rare
  select(-chdorinid_cladocera) %>% 
  I()


adonis(spp ~ treatment, data = dimorph)
  
nmds <- metaMDS(spp)  

stressplot(nmds)

fortify(nmds) %>% 
  filter(Score == "sites") %>% 
  mutate(treatment = dimorph$treatment) %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, label = Label, colour = treatment)) +
  geom_point() + 
  ggrepel::geom_text_repel()

fortify(nmds) %>% 
  filter(Score == "species") %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, label = Label)) +
  geom_point() +
  ggrepel::geom_text_repel()
```



Code to reproduce these analyses can be found at https://github.com/Sorbus-torminalis/start/blob/main/R/dimorphism.Rmd