---
title: "Ecology 2019"
author: "Sorbus torminalis"
date: "11/11/2020"
output:
  bookdown::html_document2:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r digitised-data, eval = FALSE}
library(tidyverse)
s10 <- read_csv( 
"Activity, Daphnia
0.004009048996305609, 2.705673758865248
7.051774004180864, 3.25531914893617
14.09796397583116, 3.7659574468085104
20.00873399959909, 4.156028368794327
25.08762635663354, 3.9432624113475176
31.002405429397786, 4.432624113475177
38.05647031871939, 5.138297872340425
46.03462100168953, 4.7304964539007095
56.04879585349789, 4.74822695035461
62.12050055840325, 5.124113475177305
")

#fit model
mod <- lm(log(Daphnia) ~ log(Activity + 1), data = s10) # +1 not in methods but necessary to avoid log(0)
summary(mod)# reproduces results on page 7
plot(mod)

#make predictions
nd <- tibble(Activity = 0:62)
nd$log_Daphnia <- predict(mod, newdata = nd)

#transformed
ggplot(s10, aes(x = log(Activity + 1), y = log(Daphnia))) +
  geom_point() +
  geom_line(data = nd, aes(x = log(Activity + 1), y = log_Daphnia))

#untransformed space
ggplot(s10, aes(x = Activity, y = Daphnia)) +
  geom_point() +
  geom_line(data = nd, aes(x = Activity, y = exp(log_Daphnia)))

```



Most of the data in this paper are presented as summary statistics. 
One exception is the data from the feeding trial.

> We conducted feeding trials for 10 Epitheca selected to represent a range of high and low activity rate individuals. We began by introducing each individual to a separate 4-L tank filled with filtered pond water. After allowing individuals to acclimate for 24 h, we introduced ten adult Daphnia sp., counting the number of Daphnia remaining after one hour.

Given this description, the data should consist of ten _Daphnia_ counts (integers) with the associated activity.

However, none of the ten response values in figure S10 are integers.
This plot has untransformed axes.

I digitised figure S10 and can fit a linear model that approximately reproduces the figure and reported effect size and uncertainty.

Please can the authors explain why there are non-integer counts.

One possibility is that the non-integer counts are the mean of several trials with each larvae. 
If so, this should have been described in the methods (number of replicates, how much time between trials, etc.).
However, the small difference (~0.02) in the number of larvae consumed by the larvae with activities of 46 & 56 can only be explained if an impractically large number of trial (~ 50) are used.

Another possibility is that partially devoured _Daphnia_ are counted, which would make little sense as a half-eaten _Daphnia_ is unlikely to thrive.

Another paper by the same authors reports a similar feeding trial (Ecology Letters 2017; https://doi.org/10.1111/ele.12735).
The Ecology Letters feeding trial uses the same number of the same species but a different experimental setting (Petri dish vs 4L tanks; 24 hours vs 1 hour).
It shows a very tidy linear relationship between the number of _Daphnia_ consumed (all integers) and predator activity.

Please can the authors explain why different experimental setting were used in the different papers and how this could change a linear response with no evidence of curvature to a strongly saturating response.



