---
title: "Plant sex alters Allee effects in aggregating plant parasites"
author: "Sorbus torminalis"
date: "10/12/2020"
output:
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(broom)
library(patchwork)
theme_set(theme_bw())
```

```{r download}
if(!fs::file_exists(here("data", "dryad files.xlsx"))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.c6f97")
  fs::file_move(
    path = filelocation[[1]], 
    new_path = here("data", "dryad files.xlsx")
  ) 
}
```


```{r load-data}
oikos <- readxl::read_excel(
  path =  here("data", "dryad files.xlsx"),
  sheet = "Sheet1"
)
```

The paper reports that 

> we collected all sumac aphid galls, a plant parasite, from 56 leaves (173 galls) on ﬁfteen male trees and  44 leaves (240 galls) on ﬁfteen female trees

This does not fully agree with the limited (see #2) archived data. 
Only 14 female trees are in the archived data with 43 leaves (labelled as branches in the archive). 
Conversely, all 15 male trees are in the archive data, but only 55 leaves/branches.
The reported number of galls matches the reported number.

Inspection of the data shows that tree 16 and branches 22 and 100 are missing, as are galls numbered between 80 and 88 inclusive. 
From the way the data are structured it appears that tree nr 16 was female had a single branch (nr 22) and the nine galls 80-88. 
Branch 100 was presumably on male tree nr 15 and had an unknown number of galls.

Please can the authors explain why these data have been omitted and report how this omission changed the results.


```{r}
#trees
oikos %>% 
  group_by(tree, tree.sex) %>% 
  select(-gall, -density, -branch) %>% 
  distinct() %>% 
  ungroup() %>% 
  count(tree.sex) %>% 
  pander::pander(style = "rmarkdown", split.tables = Inf, caption = "Trees")

# leaves/branches
oikos %>% 
  group_by(tree, tree.sex, branch, density) %>% 
  select(-gall) %>% 
  distinct() %>% 
  ungroup() %>% 
  count(tree.sex) %>% 
    pander::pander(style = "rmarkdown", split.tables = Inf, caption = "Branches/leaves")

#galls
oikos %>% 
  count(tree.sex) %>% 
  pander::pander(style = "rmarkdown", split.tables = Inf, caption = "Galls")

```

```{r}
setdiff(1:30, oikos$tree)
setdiff(1:100, oikos$branch)
setdiff(1:max(oikos$gall), oikos$gall)
nrow(oikos)
max(oikos$gall)
```
```{r}
oikos %>% 
  group_by(tree, branch) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  summarise(agree = mean(n ==density))

oikos %>% group_by(gall) %>% 
  count() %>% 
  filter(n > 1)
```



```{r}
oikos %>%  
ggplot( aes(x = tree, y = branch, colour = factor(tree.sex))) + geom_point() 

oikos %>% 
ggplot( aes(x = branch, y = gall, colour = factor(tree.sex))) + geom_point() 

oikos %>% 
ggplot( aes(x = gall, y = density, colour = factor(tree.sex))) + geom_point() 




```

```{r}
oikos %>% 
  ggplot(aes(x = density, fill = factor(tree.sex))) +
  geom_bar(position = position_dodge(preserve = "single"))
```


