---
title: "Predator personality structures"
author: "Sorbus torminalis"
date: "11/11/2020"
output:
  bookdown::html_document2:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(broom)
theme_set(theme_bw())
```


```{r download}
if(!fs::file_exists(here("data", "animal_personality_trophic_ecolletts2017.xlsx"))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.r58br")
  fs::file_move(
    path = filelocation[[1]], 
    new_path = here("data", "animal_personality_trophic_ecolletts2017.xlsx")
  ) 
}
```


```{r import}
personality <- readxl::read_excel(
  path = here("data", "animal_personality_trophic_ecolletts2017.xlsx"), 
  sheet = "Sheet1", na = "NA")

personality <- personality %>% 
  mutate(row = 1:n())
```

<!-- # Missing data -->

<!-- While some data are archived for this paper, much is not.  -->

<!-- I believe the following data sets were collected for and presented in this paper but have not been archived. Brackets show where the data are presented. -->

<!-- - Head widths (S1) -->
<!-- - Response to other predators (S2)  -->
<!-- - Post-experiment _Epitheca_ activity (S3) -->
<!-- - Natural densities of _Epitheca_ (S4) -->
<!-- - Initial and interim counts of zooplankton (S5), copepods (S7), and algae (S10) -->
<!-- - Timing of cannibalism (S12) -->
<!-- - Zooplankton susceptibility to predation (Fig 1) -->
<!-- - Zooplankton impact of algal abundances (Fig 1) -->

<!-- Please can the authors archive these data sets. -->

<!-- I also ask that the authors archive their data analysis code so that apparent discrepancies between the methods and the figures noted by #1 can be better understood. -->

## Non-random allocation of larvae

The paper report that 

> we randomly assigned individual predators to replicates of both predator treatments.

As #1 observes, this does not be the case as in treatment 2P the pairs of larvae either have very similar or dissimilar activities.
Plotting activity against cup number (Fig. \@ref(fig:activity-cup)) makes this pattern clearer. 
It appears that the pattern observed by #1 is not a hidden treatment but a systematic attempt to cover all parts of the activity space (low-low, low-high, high-high).
Furthermore, the regular pattern in the allocation to treatment shown by the grey line in (Fig. \@ref(fig:ranked-activity)) suggests that allocation to treatment was not done randomly random but in a systematic fashion by activity as, for example, larvae with adjacent ranks are never allocated to consecutive cups in treatment 1P

Please can the authors explain the apparent discrepancy between the description of the methods and what the data suggest was done.

```{r}
# personality %>% 
#   filter(treatment == "2P") %>% 
#   ggplot(aes(x = total.dist.one,  y = total.dist.two)) +
#   geom_abline(colour = "grey45") +
#   geom_point() 

# personality %>% 
#   mutate(cannibalism = factor(cannibalism), 
#          cannibalism = fct_explicit_na(cannibalism, na_level = "-")) %>% 
# ggplot(aes(x = row,  y = final.plankton, colour = treatment, shape = cannibalism)) +
#   geom_point()
```

```{r activity-cup, fig.cap = "Activity against cup"}
personality %>% 
  select(cup, total.dist.one, total.dist.two, treatment) %>% 
  pivot_longer(starts_with("total")) %>% 
  filter(!is.na(value)) %>% 
  ggplot(aes(x = cup, y = value,  colour = name)) + 
  geom_point() + 
  facet_wrap(~ treatment, scales = "free_x") +
  labs(x = "Cup number", y = "Activity")
 
```
 
 
```{r ranked-activity, fig.cap="Ranked activity ploted against cup number. Grey lines joins larvae by rank. Colours show when multiple larvae have the same activity."} 
# personality %>% 
#   filter(str_detect(treatment, "P")) %>% 
#   select(cup, treatment, total.dist.one, total.dist.two) %>% 
#   pivot_longer(starts_with("total")) %>% 
#   filter(!is.na(value)) %>% 
#   mutate(rank = rank(value, ties.method = "min")) %>% 
#   ggplot(aes(x = cup, y = rank, shape = treatment, colour = name)) + 
#   geom_point() 

personality %>% 
  filter(str_detect(treatment, "P")) %>% 
  select(cup, treatment, total.dist.one, total.dist.two) %>% 
  pivot_longer(starts_with("total")) %>% 
  filter(!is.na(value)) %>% 
  mutate(rank = rank(value, ties.method = "min")) %>% 
  select(-value) %>% 
  group_by(rank) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  mutate(n = factor(n)) %>% 
  arrange(desc(rank)) %>% 
  ggplot(aes(x = cup, y = rank, colour = n, shape = treatment)) + 
  geom_point() +
  geom_path(group = 1, colour = "grey40") +
  labs(x = "Cup number", y = "Rank activity")
```

## Last digits of the response variables

The distribution of the digits of the response variables should be approximately uniform, but they are not for final.plankton or final algae (Fig. \@ref(fig:last-digit)).

```{r last-digit, fig.cap = "Distribution of the last digit of the response variables. Dashed red line shows the expected value.", message = FALSE}
personality_last <- personality %>% 
  select(treatment, starts_with("final")) %>% 
  pivot_longer(-treatment) %>% 
  filter(!(treatment == "A" & name %in% c("final.plankton", "final.cop"
))) %>% #experimentally set to zero
  mutate(last = value %% 10) %>% 
  count(name, last)

ggplot(personality_last, aes(x = last, y = n)) +
  geom_col(orientation = "x") +
  geom_hline(aes(yintercept = expected), data = personality_last %>% group_by(name) %>% summarise(expected = sum(n)/10), colour = "red", linetype = "dashed") +
  facet_wrap( ~ name, ncol = 1) +
  scale_x_continuous(breaks = 0:9) +
  labs(x = "Last digit", y = "Frequency")

personality_last %>%
  group_by(name) %>% 
  nest() %>% 
  mutate(chi = map(data, ~chisq.test(.x$n)),
         chi2 = map(chi, glance)) %>% 
  select(-data, -chi) %>% 
  unnest(cols = chi2) %>% 
  select(name, p.value) %>% 
  pander::pander(style = "rmarkdown", caption = "Chi square analysis of the last digits")

# personality_last %>% 
#   pivot_wider(names_from = "name", values_from = "n") %>% 
#   select(-last) %>% 
#   cor()
```

One possible explanation is that the counts are approximate for the final algae and final plankton, but that the counts are exact for the copepods which have much lower counts.

Another paper by the same authors also has an uneven distribution of last digits with few threes. 
https://pubpeer.com/publications/43DF36777257259B8F6FED576F96F3#4

## Method confusion

As #1 has pointed out, there is confusion as to what sort of model was fitted in the models predicting the zooplankton and algal abundances from activity rate and treatment. 
The figures (2b, 3b, 4b) all show straight lines, indicative of ordinary least squares, but the captions report that 

> Best fit lines are predicted values from a GLM.

This could actually be true is the Gaussian family is chosen (or used by default if the `family` argument to `glm` is not set).

The methods' section is unclear what was done. 
First it reports:

> We used LMs to test the effects of predator personality on zooplankton abundance, community structure and algal abundance in mesocosms. 

Later in the same paragraph, it appears to contradict this:

>We first used a Poisson GLM to test for differences in total prey abundance, including treatment and activity rate as main effects.

A Poisson GLM is also reported the algal abundances, and the binomial GLM for the proportion copepod.

Finally, the paper reports that:

> Statistical analyses were conducted in R (...) using the base and ‘lme4’ packages (...).

It is not clear why the `lme4` package was used as none of the analyses appear to use a mixed effect model (nor do mixed effect models appear to be appropriate).

I can replicate the figures showing the regression lines and some of the numerical results (for example p-value of interaction of activity and treatment on zooplankton abundance) using `lm`.
For other numerical results, there is insufficient information. 

Please can the authors clarify which methods they used for these analyses, justify their decision, and archive the code used.

### Figure 2a, 3a, and 4a

Figure 2a and 4a show the abundance of zooplankton and algae, respectively, and figure 3a shows the relative abundance of copepods. 
The figure captions all report that 

> Error bars represent 1 SD

The error bars are not symmetrical about the central value for each treatment (
from the raw data, these central values appears to be the means).

The error bars are not 1 SD, as this would give wider and symmetrical errors. 
Nor are they the standard errors of a Poisson model, as these would be narrower (and almost symmetrical) (Fig \@ref(fig:fig-2a)).

```{r fig-2a, fig.cap = "Central point and error bars either digitised from the figure 2a or calculated in different ways for treatment 1P. Raw data with violin plot showing quartiles also shown.", message = FALSE}
digitised <- tibble(
  lower = 75.7396449704142, 
  central = 81.65680473372781, 
  upper =  94.67455621301775)

#poisson model
mod2a <- personality %>% 
  filter(treatment != "A") %>% 
  glm(final.plankton ~ treatment, data = ., family = "poisson") %>% 
  predict(newdata = tibble(treatment = c("1P", "2P", "D")), type = "link", se.fit = TRUE)

ex <- 1
mod2a_pred <- tibble(
  treatment = c("1P", "2P", "D"), 
  central = mod2a$fit,
  upper = central + mod2a$se.fit * ex,
  lower = central - mod2a$se.fit * ex) %>% 
  mutate(across(-treatment, exp))


# mean sd
mean_sd <- personality %>% 
  filter(treatment != "A") %>% 
  group_by(treatment) %>% 
  summarise(central = mean(final.plankton), med = median(final.plankton), sd = sd(final.plankton), lower = central - sd, upper = central + sd)

# mean sd
skewball <- personality %>% 
  filter(treatment != "A") %>% 
  group_by(treatment) %>% 
  summarise(central = mean(final.plankton), sd = sd(final.plankton), lower = central - 0.25 * sd, upper = central + 0.5 * sd)


bind_rows(Digitised = digitised,
  #        "Mean +/- sd" = mean_sd %>% filter(treatment == "1P"), 
          "Mean +0.8/-0.2 sd" = skewball %>% filter(treatment == "1P"),
          "Poisson glm +/- se" = mod2a_pred %>% filter(treatment == "1P"),
          .id = "method") %>% 
  ggplot(aes(x = method, y = central, ymin = lower, ymax = upper)) +
  geom_pointrange() +
    geom_violin(aes(x = "Raw", y = final.plankton),
                data = personality %>% filter(treatment == "1P"), draw_quantiles = (0.25 * 1:3), inherit.aes = FALSE) +
  geom_jitter(aes(x = "Raw", y = final.plankton),
                data = personality %>% filter(treatment == "1P"), inherit.aes = FALSE) 
```

Please can the authors state how they calculated the error bars on these plots.


```{r replicate, eval = FALSE}
#fig2b

personality %>% filter(str_detect(treatment, "P")) %>% 
  ggplot(aes(x = mean.total.dist, y = final.plankton, colour = treatment)) +
    geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(method = "glm", method.args = list(family = poisson), linetype = "dashed", se = FALSE)

```

```{r, eval = FALSE}
#fig3

personality %>% filter(treatment != "A") %>%
  ggplot(aes(x = treatment, y = prop.cop)) +
    geom_violin() +
  geom_jitter() +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), colour = "red") 

personality %>% filter(str_detect(treatment, "P")) %>% 
  ggplot(aes(x = mean.total.dist, y = prop.cop, colour = treatment, weight = final.plankton)) +
    geom_point() +
  geom_smooth(method = "lm") +
  geom_smooth(method = "glm", method.args = list(family = binomial), linetype = "dashed")
```

```{r, eval = FALSE}
#fig4

personality %>%
  ggplot(aes(x = treatment, y = final.algae)) +
    geom_violin() +
  geom_jitter() +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), colour = "red") 

personality %>% filter(str_detect(treatment, "P")) %>% 
  ggplot(aes(x = mean.total.dist, y = final.algae, colour = treatment)) +
    geom_point() +
  geom_smooth(method = "lm") 

##fig 5
personality %>% filter(str_detect(treatment, "P")) %>%
  ggplot(aes(x = prop.cop, y = final.algae, colour = treatment)) + 
  geom_point() +
   geom_smooth(method = "lm", aes(group = 1)) 

personality %>% filter(str_detect(treatment, "P")) %>%
  ggplot(aes(x = final.plankton, y = final.algae, colour = treatment)) + geom_point() +
   geom_smooth(method = "lm", aes(group = 1)) +
  geom_smooth(method = "glm", aes(group = 1), method.args = list(family = poisson), colour = "red")
```


```{r eval = FALSE}

#2b
personality %>% 
  filter(treatment %in% c("1P", "2P")) %>% 
lm(final.plankton ~  mean.total.dist * treatment, data = .) %>% 
  summary()

personality %>% 
  filter(treatment %in% c("1P", "2P")) %>% 
glm(final.plankton ~  mean.total.dist * treatment, data = ., family = poisson) %>% 
  summary()

#no cannibals
personality %>% 
  filter(treatment %in% c("1P", "2P"), is.na(cannibalism) | cannibalism == 0) %>% 
lm(final.plankton ~  mean.total.dist * treatment, data = .) %>% 
  summary()

personality %>% 
  filter(treatment %in% c("1P", "2P"),  is.na(cannibalism) | cannibalism == 0) %>% 
glm(final.plankton ~  mean.total.dist * treatment, data = ., family = poisson) %>% 
  summary()


personality %>%
  filter(treatment == "2P") %>% 
  ggplot(aes(x = factor(cannibalism), y = prop.cop)) + geom_point() +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), colour = "red") 

```

### Supplementary analyses

Appendix B reports an an analysis of the effect of the difference in predator activity on the response variables.
A `gls` is used to account for unequal variance. 
Unfortunately, it is not clear from the text which variance function was used nor how it was parametrised.
Neither is it clear which variables were used as predictors in the model (in addition to the difference between activities, it would seem sensible to include mean activity as it has already been shown to have a large effect).
As a result of these omissions, I have been unable to replicate the results in appendix B.

```{r appendix-B, eval = FALSE}
library(nlme)
library(broom.mixed)

personality %>% 
  filter(treatment == "2P") %>% 
  mutate(delta = total.dist.two - total.dist.one) %>% 
  ggplot(aes(x = delta, y = final.plankton, colour = factor(cannibalism))) +
  geom_point() +
  geom_smooth(method = "lm", aes(group = 1), se = FALSE)


personality %>% 
  filter(treatment == "2P") %>% 
  mutate(delta = total.dist.two - total.dist.one) %>%
  select(mean.total.dist, delta, cannibalism, final.plankton, prop.cop, final.algae) %>% 
  pivot_longer(c(final.plankton, prop.cop, final.algae)) %>% 
  group_by(name) %>% 
  nest() %>% 
  mutate(gls = map(data, ~gls(value ~ delta + mean.total.dist, # predictors unknown  
                              data = .x, 
                              weights = varFixed(~ mean.total.dist))),# variance structure unknown
         res = map(gls, tidy)
         ) %>%
  unnest(res) %>% 
  select(-data, -gls) %>% 
  filter(term != "(Intercept)")
  
```

## Code for reproduction

Code to reproduce these analyses can be found at https://github.com/Sorbus-torminalis/start/blob/main/R/Predator_personality_structures.Rmd



# Extra error bars

```{r fig-2a2, fig.cap = "Central point and error bars either digitised from the figure 2a or calculated in different ways for treatment 1P. Raw data with violin plot showing quartiles also shown.", message = FALSE}
digitised <- tibble(
  treatment = "1P", name = "final.plankton",
  lower = 75.7396449704142, 
  central = 81.65680473372781, 
  upper =  94.67455621301775) %>% 
  mutate(method = "digitised")

fp <- read_csv(col_types = "-d",
"x,value  
Bar0, 74.99999999999999
Bar1, 81.32530120481925
Bar2, 95.18072289156625
Bar3, 45.18072289156624
Bar4, 51.20481927710843
Bar5, 62.048192771084324
Bar6, 140.3614457831325
Bar7, 145.18072289156626
Bar8, 154.21686746987947") %>% 
  mutate(what = rep(c("lower", "central", "upper"), times = 3), 
         treatment = rep(c("1P", "2P", "D"), each = 3))
  
pc <- read_csv(col_types = "-d",
"x,value  
Bar0, 0.1219135802469136
Bar1, 0.154320987654321
Bar2, 0.2191358024691358
Bar3, 0.279320987654321
Bar4, 0.32407407407407407
Bar5, 0.412037037037037"
) %>% 
  mutate(what = rep(c("lower", "central", "upper"), times = 2), 
         treatment = rep(c("1P", "2P"), each = 3))

aa <- read_csv(col_types = "-d", comment = "#",
"x,value  
Bar0, 43.91167192429022
Bar1, 51.48264984227129
Bar2, 65.61514195583595
Bar3, 68.13880126182964
Bar4, 73.69085173501577
Bar5, 84.79495268138803
Bar6, NA #lower range within point
Bar7, 16.65615141955837
Bar8, 20.189274447949522
Bar9, 119.11671924290219
Bar10, 124.66876971608832
Bar11, 136.27760252365928") %>% 
  mutate(what = rep(c("lower", "central", "upper"), times = 4), 
         treatment = rep(c("1P", "2P", "D", "A"), each = 3))



dig <- bind_rows(
  final.plankton = fp,
  prop.cop = pc,
  final.algae = aa,
  .id = "name"
) %>% 
  pivot_wider(names_from = "what", values_from = "value") %>% 
  mutate(method = "dig")


# mean sd
skew = c(0.25, 0.5)
mean_sd <- personality %>% 
  select(treatment, final.plankton, prop.cop, final.algae) %>% 
  pivot_longer(c(final.plankton, prop.cop, final.algae)) %>% 
  filter(!(treatment == "A" & name %in% c("final.plankton", "prop.cop"))) %>% 
  group_by(treatment, name) %>% 
  summarise(central = mean(value), 
            sd = sd(value))

mean_sd_skew <- bind_rows(
  mean_sd = mean_sd %>% mutate( 
            lower = central - sd, 
            upper = central + sd),
  skewed = mean_sd %>% mutate( 
            lower = central - skew[1] * sd, 
            upper = central + skew[2] * sd),
  dig = dig,
  .id = "method") %>% 
  mutate(method = factor(method, levels = c("mean_sd","skewed", "dig")))


mean_sd_skew %>% 
  ggplot(aes(x = method, y = central, ymin = lower, ymax = upper)) +
  geom_pointrange() +
  facet_wrap(~ paste(name, treatment), scale = "free_y")


#for lim = upper or lower. Plot |central - lim| against sd with abline slopes of 0.1 - 1
#prop cop on different plot as different scale

dig %>% 
  full_join(mean_sd %>% select(-central), by = c("name", "treatment")) %>% 
  mutate(low = central - lower,
         high = upper - central,
         type = if_else(str_detect(name, "final"), "final.plankton & final.algae", "prop.cop")) %>% 
  select(-c(lower, central, upper, method)) %>% 
  pivot_longer(c(low, high), names_to = "lowhigh", values_to = "delta") %>% 
  ggplot(aes(x = sd, y = delta, colour = name, fill = name, shape = lowhigh)) +
  geom_point() +
  geom_abline(aes(slope = slope, intercept = 0, linetype = fct_rev(factor(slope))), tibble(slope = c(0.26, 0.53)), show.legend = TRUE) +
  facet_wrap(~ type, scales = "free") +
  labs(fill = "Variable", colour = "Variable", linetype = "Slope", shape = "Upper/Lower", x = "SD from data", y = "Digitised 'SD'") +
  scale_shape_manual(values = c(24, 25)) +
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  guides(shape = guide_legend(override.aes = list(linetype = 0)), 
         fill = guide_legend(override.aes = list(linetype = 0)))



```




