---
title: "Untitled"
author: "Sorbus torminalis"
date: "11/11/2020"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
theme_set(theme_bw())
```


```{r download}
if(!fs::file_exists(here("data", "animal_personality_trophic_ecolletts2017.xlsx"))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.r58br")
  fs::file_move(
    path = filelocation[[1]], 
    new_path = here("data", "animal_personality_trophic_ecolletts2017.xlsx")
  ) 
}
```


```{r import}
personality <- readxl::read_excel(
  path = here("data", "animal_personality_trophic_ecolletts2017.xlsx"), 
  sheet = "Sheet1", na = "NA")

personality <- personality %>% 
  mutate(row = 1:n()) %>% 
  mutate(hidden_treatment = if_else(total.dist.two - total.dist.one > 5, "high2", "similar"),
         hidden_treatment = factor(hidden_treatment), 
         hidden_treatment = fct_explicit_na(hidden_treatment, na_level = "-"))
```

# Missing data

Much of the data needed to reproduce the figures in the supplementary material is not archived.

- Head width data (S1)
- Response to other predators (S2) ?
- Post-experiment _Epitheca_ activity (S3)
- Natural densities of _Epitheca_ (S4)
- Interim counts of zooplankton (S5), copepods (S7), and algae (S10)
- Timing of cannibalism (S12)
- Zooplankton susceptibility to predation (Fig 1)
- Zooplankton impact of algal abundances (Fig 1)

Please can the authors archive these data.



```{r}
personality %>% group_by(cup) %>% filter(n() > 1)

personality %>% ggplot(aes(x = row, y = cup)) + geom_point()


ggplot(personality, aes(x = total.dist.one,  y = total.dist.two, colour = hidden_treatment)) +
  geom_abline(colour = "grey45") +
  geom_point() 

personality %>% filter(total.dist.one >= total.dist.two)



ggplot(personality, aes(x = cup,  y = final.plankton, colour = treatment, shape = hidden_treatment)) +
  geom_point()

ggplot(personality, aes(x = row,  y = final.plankton, colour = treatment, shape = hidden_treatment)) +
  geom_point()

ggplot(personality, aes(x =total.dist.one,  y = final.plankton, colour = treatment)) +
  geom_point()

personality %>% filter(hidden_treatment == "high2")
personality %>% ggplot(aes(x = cup, y = individual.two, colour = hidden_treatment)) + geom_point()

personality %>% 
  select(final.plankton) %>% 
 # count(final.plankton) %>% filter(n > 2)
  ggplot(aes(x = final.plankton)) + geom_bar()


personality %>% 
  ggplot(aes(x = cup, y = final.cop, colour = treatment, shape = hidden_treatment)) + geom_point()

personality %>% 
  ggplot(aes(x = cup, y = final.algae, colour = treatment)) + geom_point()

personality %>% 
  ggplot(aes(x = final.algae, fill = treatment)) + geom_bar()




personality %>% 
  select(treatment, starts_with("final")) %>% 
  pivot_longer(-treatment) %>% 
  filter(!(treatment == "A" & name %in% c("final.plankton", "final.cop"
))) %>% #experimentally set to zero
  mutate(last = value %% 10) %>% 
  ggplot(aes(x = last)) +
  geom_bar() +
  facet_wrap( ~ name, ncol = 1) +
  scale_x_continuous(breaks = 0:9) +
  theme_minimal()

library(magrittr)
personality %>% 
  transmute(last = final.plankton %% 10) %$% chisq.test(x = last)


personality %>% 
  ggplot(aes(x = treatment, y = total.dist.one)) + geom_boxplot() # expected if large indic


personality %>% 
  select(starts_with("total.dist.")) %>% 
  pivot_longer(everything()) %>% 
  filter(!is.na(value)) %>% 
  mutate(last = value %% 10) %>% 
  filter(value > 0) %>% 
  ggplot(aes(x = last, fill = name)) + geom_bar() +  scale_x_continuous(breaks = 0:9) +
  theme_minimal()

personality %>% 
  filter(treatment %in% c("1P", "2P")) %>% 
  ggplot(aes(x = mean.total.dist, y = prop.cop, colour = hidden_treatment)) + geom_point()

personality %>% filter(hidden_treatment == "similar") %>% 
  ggplot(aes(x = individual.two, total.dist.two - total.dist.one)) + geom_point()

#fig2
personality %>% filter(treatment != "A") %>% 
  ggplot(aes(x = treatment, y = final.plankton)) +
    geom_violin() +
  geom_jitter() +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), colour = "red") 
personality %>% filter(str_detect(treatment, "P")) %>% 
  ggplot(aes(x = mean.total.dist, y = final.plankton, colour = treatment)) +
    geom_point() +
  geom_smooth(method = "lm") 

#fig3

personality %>% filter(treatment != "A") %>%
  ggplot(aes(x = treatment, y = prop.cop)) +
    geom_violin() +
  geom_jitter() +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), colour = "red") 
personality %>% filter(str_detect(treatment, "P")) %>% 
  ggplot(aes(x = mean.total.dist, y = prop.cop, colour = treatment)) +
    geom_point() +
  geom_smooth(method = "lm") 

#fig4

personality %>%
  ggplot(aes(x = treatment, y = final.algae)) +
    geom_violin() +
  geom_jitter() +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), colour = "red") 
personality %>% filter(str_detect(treatment, "P")) %>% 
  ggplot(aes(x = mean.total.dist, y = final.algae, colour = treatment)) +
    geom_point() +
  geom_smooth(method = "lm") 

##fig 5
personality %>% filter(str_detect(treatment, "P")) %>%
  ggplot(aes(x = prop.cop, y = final.algae, colour = treatment)) + geom_point() +
   geom_smooth(method = "lm", aes(group = 1)) 

personality %>% filter(str_detect(treatment, "P")) %>%
  ggplot(aes(x = final.plankton, y = final.algae, colour = treatment)) + geom_point() +
   geom_smooth(method = "lm", aes(group = 1)) +
  geom_smooth(method = "glm", aes(group = 1), method.args = list(family = poisson), colour = "red")

```



